```{r}
#| include: false
knitr::opts_chunk$set(
  echo = FALSE
)
```

```{r}
#| echo: false
cat_ascii <- "


  /\\_/\\
 ( o.o )
 > ^ <
 
"

gatai_ascii <- "

  /\\_/\\
 ( o.o )
  > ^ <
 
"

tree_ascii <- "
    /\\
   //\\\\
  ///\\\\\\
 ////\\\\\\\\
/////\\\\\\\\\\
   [][]
"
```

```{r}
#| label: fig-gatai
#| fig-cap: "GatAI the AI"
#| layout-ncol: 2
#| echo: false
cat(gatai_ascii)
cat("\n\n\nGatai, the AI\nOne smart cat")
```

# Preface {.unnumbered}

In August of 2023 I undertook to do some exploratory research on video-games available on the gato.us website. The purpose of this exercise was to better understand the potential problem space and the application of artificial intelligence (AI) to some of its extent. This volume is a write-up of some of the results and experiences, and furthermore a chronicle of the process I followed.

```{r}
#| label: x-mas-3-ascii
#| layout-ncol: 2
#| echo: false
cat("\n\nIn the Red Corner:\nOur Nemesis\nX-Mas-3")
cat(tree_ascii)
```

```{r}
#| echo: false
#| label: fig-xmas3
#| fig-cap: "X-Mas-3"

```
